An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-24 Revision/response.tex

\textbf{WIT-Real dataset:} We use the \texttt{clip-vit-base-patch32} model to encode the \texttt{page\_title} field from the WIT dataset into vectors, and extract the language type and the length of the \texttt{context\_page\_}
\texttt{description} field as real attributes, applying them in both attribute filtering and range filtering experiments.

\textcolor{blue}{
	"Additionally, we introduced two datasets with real attributes for a more comprehensive evaluation of various attribute filtering and range query algorithms: The YT-Audio-Real dataset leverages URLs provided by the YouTube8M dataset to crawl corresponding video category IDs and view counts as its real attributes; whereas for the WIT-Real dataset, we used the \texttt{clip-vit-base-patch32} model to encode the \texttt{page\_title} field into vectors, and used the language type and the length of the \texttt{context\_page\_description} field as real attributes. It should be noted that, during range filtering, due to the limitations of some algorithms, we uniformly sorted the vectors according to the query attribute."
}
%%%%%%%%%%%%%%%%%%%%%% R5 %%%%%%%%%%%%%%%%%%%%%%

\myfbox{
	\begin{metarevision}
		\textit{Evaluate if the server setting affects the relative ordering of methods in experiments}
	\end{metarevision}
}

\noindent
\textbf{Response:} 
That is a great suggestion. In order to figure out if the server setting affects the relative ordering of methods, we conducted additional experiments on a server equipped with two Intel(R) Xeon(R) Gold 6330 CPUs @ 2.00GHz using the YT-Audio-Real dataset.
The results show that the performance differences across algorithms remain small, and their relative rankings do not change significantly across servers.
We did observe slightly lower QPS on this server compared to our original one, which is mainly due to the lower CPU frequency of the Xeon 6330 processors. \textbf{(Please refer to the blue highlighted parts in Section 4.2.4 and Section 4.3.4 under Different Hardware Environments.)}

Attribute filtering:
\textcolor{blue}{"To investigate the impact of different server configurations on the performance of the algorithms, we conduct additional experiments on a server equipped with dual Intel(R) Xeon(R) Gold 6330 CPUs @ 2.00GHz using the YT-Audio-Real dataset, and the results are shown in Figure 12b. Compared with the experimental results on the original server, the relative ranking of the algorithms remains largely unchanged, indicating that the performance of these methods is not significantly affected by differences in CPU configurations. Due to the lower CPU frequency and smaller RAM capacity, the QPS values are slightly lower across all methods."}

Range filtering:
\textcolor{blue}{"Consistent with Section 4.2.4, we conduct additional range filtering experiments on the YT-Audio-Real dataset using the same server configuration, with results presented in Figure 16b. When compared with the experimental results of the original server, our findings remain consistent with those reported in Section 4.2.4. For brevity and due to space limitations, we omit further discussion of these redundant results."}

@@

%Thanks to the reviewer for suggesting that there is no explanation in Figure 7 why Vamana is not included. Figure 7 is an experiment on multi-label construction and search, and the Boolean logic of multi-label search is AND. The two algorithms using Vamana, FilteredVamana and StitchedVamana, only support Boolean logic OR for multi-label search, so we did not include them. We will add an explanation of this in the paper.

%All the evaluated algorithms in our experiments build the index in a one-shot manner, without applying any partitioning or merging strategies.
%Additionally, although some of the methods we used—such as FAISS, PUCK, DiskANN, and Milvus—support vector quantization during index construction, we do not enable quantization in any of our experiments. Therefore, the reported index build times do not include quantization.

%%%%%%%%%%%%%%%%%%%%%% D3 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{detailed}
		\textit{The paper lists AF algos (table 1) and RF algos (table 2). It would be good to list the systems in a table with their algos and capabilities (AF, RF, Graph, IVF, ...). Would make it easier to follow the experimental comparison. E.g., Fig 12 does not include Vamana/DiskANN. Should I assume that RF is not possible with Vamana, or just not implemented?}
	\end{detailed}
}

\noindent
\textbf{Response:} 
Thank you for the suggestion. We realized that the original wording is somewhat vague and led to confusion among readers. To address this, we merged the tables for attribute filtering (AF) and range filtering (RF), and clarified the support of each algorithm for AF and RF in the unified table.

### 2025-07-24 Revision/response.tex

%%%%%%%%%%%%%%%%%%%%%% R2 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{Run experiments with multi-modal queries (e.g., combining text and images) or more complex query types involving both attribute and range filters.}
	\end{metarevision}
}

\noindent
\textbf{Response:} We believe this is an excellent suggestion. In fact, in the initial version of our paper, we used the text-to-image dataset that has been widely adopted by existing multimodal ANN algorithms. In this dataset, the base vectors are images, while the query set consists of texts, resulting in a modality gap between the two. Achieving good performance on such a dataset often requires the algorithm to demonstrate stronger robustness. \textbf{(Please refer to the last paragraph of Section 4.2.3 and the green highlighted part in Section 4.3.3.)}

To further enrich the evaluation of complex multimodal queries, we add new experiments in the Multi-modal Queries sections of Section 4.2.4 and Section 4.3.4. Specifically, we construct a more challenging multimodal dataset by selecting 500,000 text vectors and 500,000 image vectors as the base set, and 5,000 text vectors and 5,000 image vectors as the query set. We then perform both attribute filtering and range queries over this dataset, combining text and image queries to further analyze the robustness of the algorithms. The corresponding experimental results are presented in Section 4.2.4 and Section 4.3.4 under Multi-modal Queries.

As for complex query types that involve both attribute filtering and range queries simultaneously, to the best of our knowledge, existing algorithms—except for general-purpose systems such as Milvus and Faiss—do not support this functionality. For example, ACORN needs to build separate indexes for each scenario and cannot handle attribute filtering and range queries simultaneously. Therefore, we do not consider this type of complex query in our experiments but instead study the two query types separately. \textbf{(Please refer to the blue highlighted parts in  Section 4.1.1 paragraph 4 and  Section 4.2.4 and Section 4.3.4 under Multi-modal Queries.)
}

\textcolor{blue}{
"Additionally, we use the Out-of-Distribution (OOD) dataset Text2Image and the large-scale dataset Deep100M to evaluate the generalization capability of the algorithms. The Text2Image dataset includes two versions: the original version contains 1 million image embeddings as the base dataset and uses 10 thousand text embeddings as the query set; the Text2Image-Mix version is composed of 0.5 million image embeddings and 0.5 million text embeddings as the base dataset, with a query set of 5 thousand image embeddings and 5 thousand text embeddings."
}

Attribute filtering: \textcolor{blue}{
	"We conducted experiments on the text2image-mix dataset. This dataset contains both image and text vectors. It is a more complex multimodal dataset. As shown in Figure 12, although the queries are more complex, UNG and NHQ still perform the best. CAPS is the next best.
	Since the dataset mixes two modalities, the ground truth of the query set usually corresponds to the same modality. The nearest neighbors in the index are often from the same modality because they are closer in distance. Therefore, the search results on this dataset are similar to those in Figure 4. UNG and NHQ perform the best. CAPS clusters based on vector similarity. It actually clusters the vectors into the same modality. Then it uses attribute filtering to divide the data further. This allows CAPS to quickly find nearest neighbors that meet the filter conditions.
	In contrast, the text2image dataset shown in Figure 11 is different. The queries and the dataset are from different modalities. The queries are far from the dataset. The ground truths are also scattered and not clustered together. To find the ground truths, the algorithm often needs to traverse more points. This is a greater challenge for the algorithms." 
}

Range filtering:\textcolor{blue}{
		"Similar to attribute filtering, the performance of multimodal queries with range filtering shown in Figure 16a is also consistent with the single-modality results in Figure 14. As shown in Figure 15, the algorithms still perform the worst on multimodal datasets like Text2Image. This is because these range query algorithms are almost graph-based. Graph algorithms need more iterations to find true neighbors in OOD datasets."
}

### 2025-07-24 VLDB2026/main.tex

	\textit{\textbf{\textcolor{blue}{Different Hardware Environments.}}} 
%	\textcolor{cyan}{To investigate the impact of different server configurations on the performance of the algorithms, we conducted additional experiments on Server B using the YT-Audio-Real dataset, and the results are shown in Figure~\ref{fig:Different Computer}. Compared with the experimental results on Server A shown in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}f, the relative ranking of the algorithms remains largely unchanged, indicating that the performance of these methods is not significantly affected by differences in CPU configurations. Due to the lower CPU frequency and smaller RAM capacity of Server B compared to Server A, the QPS values are slightly lower across all methods.}
	\textcolor{red}{To investigate the impact of different server configurations on the performance of the algorithms, we conduct additional experiments on a server equipped with dual Intel(R) Xeon(R) Gold 6330 CPUs @ 2.00GHz using the YT-Audio-Real dataset, and the results are shown in Figure~\ref{fig:Different Computer}. Compared with the experimental results on the original server (as shown in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}(f)), the relative ranking of the algorithms remains largely unchanged, indicating that the performance of these methods is not significantly affected by differences in CPU configurations. Due to the lower CPU frequency and smaller RAM capacity, the QPS values are slightly lower across all methods.}

	\textit{\textbf{\textcolor{blue}{Large-scale Datasets.}}} 
	\textcolor{violet}{As shown in the figure~\ref{fig:Big Dataset}, On the 100M dataset, UNG and NHQ still exhibit excellent performance, demonstrating their good adaptability to large-scale datasets; however, this also comes with significant memory overhead. The clustering-based algorithm Faiss+HQI\_Batch shows moderate performance but has very low memory overhead, which makes it highly attractive for practical applications. We did not present the experimental results for StitchedVamana because we found that its recall consistently remained at a very low level. Its search starting point is set to 0 by default, and when the entry point is far from the target area, it easily gets stuck in a local optimum, making it difficult to navigate to the vicinity of the correct point.}

@@

	\textit{\textbf{\textcolor{blue}{Different Hardware Environments.}}}
	\textcolor{cyan}{Consistent with Section 4.2.4, we conduct additional range filtering experiments on the YT-Audio-Real dataset using the same server configuration, with results presented in Figure~\ref{fig:Different_Computer_range}. When compared with the experimental results of the original server shown in Figure~\ref{fig:exp_8_2}a, our findings remain consistent with those reported in Section 4.2.4. For brevity and due to space limitations, we omit further discussion of these redundant results.}
%	\textcolor{cyan}{Consistent with Section 4.2.4, we conducted additional range filtering experiments using the YT-Audio-Real dataset on the same server, with results shown in Figure \ref{fig:Different Computer range}. Compared with the original server's experimental results (Figure \ref{fig:exp_8_2}($e$)), the conclusions remain identical to those described in Section 4.2.4. Due to space constraints, these details are not reiterated here.}

	\textit{\textbf{\textcolor{red}{Large-scale Datasets.}}} \textcolor{red}{As shown in the figure~\ref{fig:Big_Dataset_range}, On the 100M dataset, the top 3 ranked algorithms (WinFilter, UNIFY, and iRange) remain unchanged. However, their high performance is achieved at the cost of extremely high memory consumption, as detailed in Figure~\ref{fig:rangeFilter_search_memory_mb}. Meanwhile, VBASE and Faiss demonstrate moderate performance but significantly lower memory consumption than the top three algorithms}.

	\textit{\textbf{\textcolor{blue}{Multi-modal Queries}}}
	\textcolor{blue}{
	Similar to attribute filtering, the performance of multimodal queries with range filtering shown in Figure~\ref{fig:MultiModal_range} is also consistent with the single-modality results in Figure~\ref{fig:exp_8_2}. As shown in Figure~\ref{fig:exp_8_3}, the algorithms still perform the worst on multimodal datasets like Text2Image. This is because these range query algorithms are almost graph-based. Graph algorithms need more iterations to find true neighbors in OOD datasets.}

%	\textit{\textbf{Multimodel, 85 and Deep100M.}} \textcolor{red}{The fig is Figure \ref{fig:TrinityExtension——Range}}

	% Legend
	\includegraphics[width=0.98\columnwidth]{figures/exp/range_legend.pdf}
	\vspace{0.2cm} % Added a little space after the legend

	% --- Subfigure (a) containing two plots ---
	\begin{subfigure}[t]{0.98\columnwidth}
		\centering
		% First plot for the shared subfigure
		\includegraphics[width=0.507\linewidth]{figures/exp/range_multimodel.pdf}
		\hfill % Pushes the two plots apart
		% Second plot for the shared subfigure
		\includegraphics[width=0.48\linewidth]{figures/exp/range_multimodel_1.pdf}

		% Shared caption for the two plots above
		\captionsetup{font=small}
		\caption{MultiModal}
		\label{fig:MultiModal_range} % A single label for the combined subfigure
	\end{subfigure}

	% --- Subfigure (b) ---
	\begin{subfigure}[t]{0.507\columnwidth}
		\centering
		\includegraphics[width=\linewidth]{figures/exp/range_85.pdf}
		\captionsetup{font=small}
		\caption{Different Computer}
		\label{fig:Different_Computer_range}
	\end{subfigure}
	\hfill
	% --- Subfigure (c) ---
	\begin{subfigure}[t]{0.48\columnwidth}
		\centering
		\includegraphics[width=\linewidth]{figures/exp/range_deep100M.pdf}
		\captionsetup{font=small}
		\caption{Big Dataset}
		\label{fig:Big_Dataset_range}
	\end{subfigure}

	\caption{\textcolor{red}{TrinityExtension——Range}}
	\label{fig:TrinityExtension——Range}
\end{figure}

	\section{DISCUSSION}
	Based on the performance of the algorithms on different experimental scenarios, we discuss our findings as follows.

>>>>

	\textit{\textbf{\textcolor{blue}{Different Hardware Environments.}}} 
%	\textcolor{cyan}{To investigate the impact of different server configurations on the performance of the algorithms, we conducted additional experiments on Server B using the YT-Audio-Real dataset, and the results are shown in Figure~\ref{fig:Different Computer}. Compared with the experimental results on Server A shown in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}f, the relative ranking of the algorithms remains largely unchanged, indicating that the performance of these methods is not significantly affected by differences in CPU configurations. Due to the lower CPU frequency and smaller RAM capacity of Server B compared to Server A, the QPS values are slightly lower across all methods.}
	\textcolor{red}{To investigate the impact of different server configurations on the performance of the algorithms, we conduct additional experiments on a server equipped with dual Intel(R) Xeon(R) Gold 6330 CPUs @ 2.00GHz using the YT-Audio-Real dataset, and the results are shown in Figure~\ref{fig:attribute-cross-platform}. Compared with the experimental results on the original server, the relative ranking of the algorithms remains largely unchanged, indicating that the performance of these methods is not significantly affected by differences in CPU configurations. Due to the lower CPU frequency and smaller RAM capacity, the QPS values are slightly lower across all methods.}

	\textit{\textbf{\textcolor{blue}{Large-scale Datasets.}}} 
	\textcolor{violet}{As shown in the figure~\ref{fig:attribute big dataset}, On the 100M dataset, UNG and NHQ still exhibit excellent performance, demonstrating their good adaptability to large-scale datasets; however, this also comes with significant memory overhead. The clustering-based algorithm Faiss+HQI\_Batch shows moderate performance but has very low memory overhead, which makes it highly attractive for practical applications. We did not present the experimental results for StitchedVamana because we found that its recall consistently remained at a very low level. Its search starting point is set to 0 by default, and when the entry point is far from the target area, it easily gets stuck in a local optimum, making it difficult to navigate to the vicinity of the correct point.}

@@

	\textit{\textbf{\textcolor{blue}{Different Hardware Environments.}}}
	\textcolor{cyan}{Consistent with Section 4.2.4, we conduct additional range filtering experiments on the YT-Audio-Real dataset using the same server configuration, with results presented in Figure~\ref{fig:range-cross-platform}. When compared with the experimental results of the original server, our findings remain consistent with those reported in Section 4.2.4. For brevity and due to space limitations, we omit further discussion of these redundant results.}
%	\textcolor{cyan}{Consistent with Section 4.2.4, we conducted additional range filtering experiments using the YT-Audio-Real dataset on the same server, with results shown in Figure \ref{fig:Different Computer range}. Compared with the original server's experimental results (Figure \ref{fig:exp_8_2}($e$)), the conclusions remain identical to those described in Section 4.2.4. Due to space constraints, these details are not reiterated here.}

	\textit{\textbf{\textcolor{red}{Large-scale Datasets.}}} \textcolor{red}{As shown in the figure~\ref{fig: range big dataset}, On the 100M dataset, the top 3 ranked algorithms (WinFilter, UNIFY, and iRange) remain unchanged. However, their high performance is achieved at the cost of extremely high memory consumption, as detailed in Figure~\ref{fig:rangeFilter_search_memory_mb}. Meanwhile, VBASE and Faiss demonstrate moderate performance but significantly lower memory consumption than the top three algorithms}.

	\textit{\textbf{\textcolor{blue}{Multi-modal Queries}}}
	\textcolor{blue}{
	Similar to attribute filtering, the performance of multimodal queries with range filtering shown in Figure~\ref{fig:range-multimodal} is also consistent with the single-modality results in Figure~\ref{fig:exp_8_2}. As shown in Figure~\ref{fig:exp_8_3}, the algorithms still perform the worst on multimodal datasets like Text2Image. This is because these range query algorithms are almost graph-based. Graph algorithms need more iterations to find true neighbors in OOD datasets.}

%	\textit{\textbf{Multimodel, 85 and Deep100M.}} \textcolor{red}{The fig is Figure \ref{fig:TrinityExtension——Range}}
	\section{DISCUSSION}
	Based on the performance of the algorithms on different experimental scenarios, we discuss our findings as follows.
